---
title: "Geographic distance between Canadian Cities"
link-citations: true
output: html_notebook
abstract: This script is related to "Geographic_distance_CA_cities.csv" file in the Data folder that represents the distance between Canadian cities. I have used `Maps` R-package to extract the distances. I should tahnk you Dominic Roye for publishing an informative blog in [R-bloggers](https://www.r-bloggers.com/2019/10/geographic-distance/) and his advise. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Load packages
```{r}
library(maps) 
library(sf) 
library(tidyverse) 
library(units) 
library(rnaturalearth) 
library(rnaturalearthdata)
```
##### From the maps package(Call world cities of population greater than about 40,000)
```{r}
head(world.cities) 
```
##### Convert the points into an sf object with CRS WSG84
```{r}
cities <- st_as_sf(world.cities, coords = c("long", "lat"), crs = 4326)
cities
```
##### Filter the Canadian cities
```{r}
CA_cities <- filter(cities, country.etc == "Canada")

# create a new label combining name and country
CA_cities <- mutate(CA_cities, city_country = str_c(name, " (", country.etc, ")"))
CA_cities
```
##### Example: Calculate the distance Saint John to Toronto, Calgary and Vancouver
```{r}
dist_st.Jh <- st_distance(slice(CA_cities, 663), 
                              slice(CA_cities, c(831, 106, 852)))
dist_st.Jh# distances are in meters
```
##### Change distance from meter to kilometer
```{r}
set_units(dist_st.Jh, "km")

as.vector(dist_st.Jh) ## shows as a vector
```
##### Calculate all distance between Candian cities
```{r}
distance_all <- st_distance(CA_cities)
# matrix
dim(distance_all)
# change m to km
distance_all_km <- set_units(distance_all, km)
# replace the distance of 0 m with NA
#distance_all_km[distance_all_km == set_units(0, km)] <- NA
dim(distance_all_km)

```
##### Export the distance matrix as a csv file
```{r}
Dis_df <- as.data.frame(distance_all_km)
name_cities <- data.frame(CA_cities$name)
new <- Dis_df
colnames(new) <- name_cities[,1] 
distance_all_km_withname <- cbind(name_cities, new)

write.table(distance_all_km_withname, file = "Geographic_distance_CA_cities.csv",  quote = TRUE, sep = ", ", row.names = FALSE, col.names = TRUE, qmethod = "double")
```

##### Check the csv file is correct
```{r}
dist_test<- st_distance(slice(CA_cities, 1), 
                              slice(CA_cities, c(2,5)))
dist_test# distance in meters
set_units(dist_test, "km")

```


##### Get the index (position) of the city and the distance(nearest city)
```{r}
pos <- apply(distance_all_km, 1, which.min)
dist <- apply(distance_all_km, 1, min, na.rm = TRUE)
# add the distance and get the name of the city
CA_cities <- mutate(CA_cities, nearest_city =  city_country[pos], 
                             geometry_nearest = geometry[pos],
                             distance_city = dist)
```
##### Map
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf", country = 'canada')
# map
ggplot(world) +
   geom_sf(fill = "lightblue", colour = "white") +
   geom_sf(data = CA_cities, 
           aes(size = distance_city),
           alpha = 0.7,
           fill = "#bd0026",
           shape = 21,
           show.legend = 'point') +
   coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs") +
  labs(size = "Distance (km)", title = "Distance to the next city(Canada)") +
  theme_void()


```

